### 2018-05-15 2_wmm.tex
Reduce space between letters in notations.

A \textit{memory event} $e_m \in \mathbb{E}$ represents the fact of access to the memory.
Only memory events change the state of an abstract machine executing the code, since it is completely determined by values stored in its memory.
Since memory is the crucial low-level resource shared by multiple processes, most relations are defined over memory events. 
The processes can access a shared memory location (denoted by~$l_i$, for \textit{location}), or a local one (denoted by~$r_i$, for \textit{register}).
A memory event can access at most one shared memory location, high-level instructions that address more than one shared variable must be transformed into a sequence of events.
A memory event is specified by its direction with respect to the shared variable, its location~$\mathtt{loc}(e_m)$, its processor label~$\mathtt{proc}(e_m)$, and a unique event label~$\mathtt{id}(e_m)$~\cite{alglave2010shared}.
%\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local

The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\mathtt{load}(r_1~\leftarrow~l_2); \ e_2~=~\mathtt{store}(l_1~\leftarrow~r_1)$.
Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.

@@

\vspace{1em}
The basic relations are the following~\cite{alglave2010shared}:
%TODO: say that po and rf are  defined from the candidate execution:
%TODO: say that 
\begin{itemize}
    \item The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subseteq~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
    For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \relation{po} e_2$.
    %Some new relations may be acquired : dp, po-loc

@@

    \item the \textit{from-read} relation $\fr \subseteq~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events succeeding the write event from which the read event gets its value: \\
    $r \relation{fr} w ~\triangleq~ (\exists w' . w' \relation{rf} r \land w' \relation{co} w)$;

@@

    \item the \textit{communication} relation \com over memory events, that fully describes the data-flow of a program: \\
    %$\mathtt{com} \triangleq (\mathtt{rf} \cup \mathtt{co} \cup \mathtt{fr})$
    $m_1 \relation{com} m_2~\triangleq~((m_1\relation{rf}m_2) \lor (m_1\relation{co}m_2) \lor (m_1\relation{fr}m_2))$;

    \item the \textit{external} (and \textit{internal}) \textit{from-read} relations that restrict the \fr-relation to the different (respectively, same) processes: \\
    $w \relation{fre} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) = \mathtt{proc}(r))$, \\
    $w \relation{fri} r ~\triangleq~ (w \relation{fr} r \land \mathtt{proc}(w) \neq \mathtt{proc}(r))$;

    %TODO: say that po-relation is *static*
    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
    $m_1 \relation{po-loc} m_2 ~\triangleq~ (m_1 \relation{po} m_2 \land \mathtt{loc}(m_1) = \mathtt{loc}(m_2))$; and

@@

%An \textit{execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
%The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
%associated with the instructions of the program
%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
%An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\mathtt{rf}$ and $\mathtt{co}$~\cite{alglave2010shared}.
Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus test Example~\ref{simple_wmm_x86} (the pictures are generated by the \tool{herd7} tool, version~7.47).
Since there are no conditional jumps, the \po-relation is defined and we do not need to guess it.
Since each thread performs single write followed by a single read, the \co-relation is also defined (it relates the initial write event with the write event to the same location).

@@

\begin{enumerate}
  \item \textit{Unary operations}:
    \begin{itemize}
      \item \textit{the complement} of a relation \rel{r} is $\rel{\textasciitilde r}$,
      \item \textit{the transitive closure} of a relation \rel{r} is $\rel{r+}$,
      \item \textit{the reflexive closure} of a relation \rel{r} is $\rel{r?}$,
      \item \textit{the reflexive-transitive closure} of a relation \rel{r} is $\rel{r*}$, and
      \item \textit{the inverse} of a relation \rel{r} is $\invrel{r}$.
    \end{itemize}
  \item \textit{Binary operations:}
  \begin{itemize}
    \item \textit{the union} of two relations \rel{r1} and \rel{r2} is \rel{r1\,|\,r2},
    \item \textit{the intersection} of two relations \rel{r1} and \rel{r2} is \rel{r1\,\&\,r2},
      \item \textit{the difference} of two relations \rel{r1} and \rel{r2} is $\mathtt{r1 \relminus r2}$, and
    \item \textit{the sequence} of two relations \rel{r1} and \rel{r2} is \rel{r1;r2}, which is defined as the set of pairs \rel{(x,y)} such that there exists an intervening \rel{z}, such that $\rel{(x,z)} \in \rel{r1}$ and $\rel{(z,y)} \in \rel{r2}$.
  \end{itemize}
\end{enumerate}

For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\mathr{fr} = (\mathr{rf}^{-1};\mathr{co})$. As an example memory model definition in \cat{} language, the the x86-TSO can be found in Appendix~\ref{apx:x86cat}. This memory model asserts acyclicity of communication relation, \rel{po-loc} relation, \rel{mfence} relation and some other derived relations~\cite{owens2009better}.

>>>>

A \textit{memory event} $e_m \in \mathbb{E}$ represents the fact of access to the memory.
Only memory events change the state of an abstract machine executing the code, since it is completely determined by values stored in its memory.
Since memory is the crucial low-level resource shared by multiple processes, most relations are defined over memory events. 
The processes can access a shared memory location (denoted by~$l_i$, for \textit{location}), or a local one (denoted by~$r_i$, for \textit{register}).
A memory event can access at most one shared memory location, high-level instructions that address more than one shared variable must be transformed into a sequence of events.
A memory event is specified by its direction with respect to the shared variable, its location~$\texttt{loc}(e_m)$, its processor label~$\texttt{proc}(e_m)$, and a unique event label~$\texttt{id}(e_m)$~\cite{alglave2010shared}.
%\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local

The set of memory events $\mathbb{M}$ is divided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
We add a restriction that each memory event uses at most one shared location, so that the write instruction $i = write(l_1, l_2)$, that encodes the write from the shared location $l_2$ to the shared location $l_1$, is represented as two consequent events $e_1~=~\texttt{load}(r_1~\leftarrow~l_2); \ e_2~=~\texttt{store}(l_1~\leftarrow~r_1)$.
Also, it is important to separate the set of initial write events $\mathbb{IW}~\subseteq~\mathbb{W}$ that perform initialisation of program variables.

@@

\vspace{1em}
The basic relations are the following~\cite{alglave2010shared}:
%TODO: say that po and rf are  defined from the candidate execution:
%TODO: say that 
\begin{itemize}
    \item The \textit{control-flow} of a program is defined by the \textit{program-order} relation \po~$\subseteq~\mathbb{E}~\times~\mathbb{E}$, which represents the total order of events of same process.
    For instance, if the instruction $i_1$ generates the event $e_1$ and the instruction $i_2$ follows $i_1$ and generates the event $e_2$, then $e_1 \relarrow{po} e_2$.
    %Some new relations may be acquired : dp, po-loc

@@

    \item the \textit{from-read} relation $\fr \subseteq~\mathbb{R}~\times~\mathbb{W}$ that maps a read event to all write events succeeding the write event from which the read event gets its value: \\
    $r \relarrow{fr} w ~\triangleq~ (\exists w' . w' \relarrow{rf} r \land w' \relarrow{co} w)$;

@@

    \item the \textit{communication} relation \com over memory events, that fully describes the data-flow of a program: \\
    %$\texttt{com} \triangleq (\texttt{rf} \cup \texttt{co} \cup \texttt{fr})$
    $m_1 \relarrow{com} m_2~\triangleq~((m_1\relarrow{rf}m_2) \lor (m_1\relarrow{co}m_2) \lor (m_1\relarrow{fr}m_2))$;

    \item the \textit{external} (and \textit{internal}) \textit{from-read} relations that restrict the \fr-relation to the different (respectively, same) processes: \\
    $w \relarrow{fre} r ~\triangleq~ (w \relarrow{fr} r \land \texttt{proc}(w) = \texttt{proc}(r))$, \\
    $w \relarrow{fri} r ~\triangleq~ (w \relarrow{fr} r \land \texttt{proc}(w) \neq \texttt{proc}(r))$;

    %TODO: say that po-relation is *static*
    \item the \texttt{po-loc} relation that is the \po-relation over events that access to the same shared variable: \\
    $m_1 \relarrow{po-loc} m_2 ~\triangleq~ (m_1 \relarrow{po} m_2 \land \texttt{loc}(m_1) = \texttt{loc}(m_2))$; and

@@

%An \textit{execution} (trace, run) of a program is an ordered set of events defined by \po- and \rf-relations and set of final writes to a given memory location that is valid under certain memory model~\cite{alglave2014herding}.
%The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
%associated with the instructions of the program
%An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}.
%An execution is uniquely defined by the set $\mathbb{X}$ of events have been executed in each thread (the \textit{control-flow} of a program), and the relations $\texttt{rf}$ and $\texttt{co}$~\cite{alglave2010shared}.
Figure~\ref{simple_wmm_x86_pic} illustrates four possible candidate executions for the litmus test Example~\ref{simple_wmm_x86} (the pictures are generated by the \tool{herd7} tool, version~7.47).
Since there are no conditional jumps, the \po-relation is defined and we do not need to guess it.
Since each thread performs single write followed by a single read, the \co-relation is also defined (it relates the initial write event with the write event to the same location).

@@

\begin{enumerate}
  \item \textit{Unary operations}:
    \begin{itemize}
      \item the \textit{complement} of a relation \texttt{r} is $\texttt{\textasciitilde r}$,
      \item the \textit{transitive closure} of a relation \texttt{r} is $\texttt{r+}$,
      \item the \textit{reflexive closure} of a relation \texttt{r} is $\texttt{r?}$,
      \item the \textit{reflexive-transitive closure} of a relation \texttt{r} is $\texttt{r*}$, and
      \item the \textit{inverse} of a relation \texttt{r} is $\invrel{r}$.
    \end{itemize}
  \item \textit{Binary operations:}
  \begin{itemize}
    \item the \textit{union} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,|\,r2},
    \item the \textit{intersection} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1\,\&\,r2},
      \item the \textit{difference} of two relations \texttt{r1} and \texttt{r2} is $\texttt{r1} \relminus \texttt{r2}$, and
    \item \textit{the sequence} of two relations \texttt{r1} and \texttt{r2} is \texttt{r1;r2}, which is defined as the set of pairs \texttt{(x,y)} such that there exists an intervening \texttt{z}, such that $\texttt{(x,z)} \in \texttt{r1}$ and $\texttt{(z,y)} \in \texttt{r2}$.
  \end{itemize}
\end{enumerate}

For instance, the \fr-relation is defined as a sequence of inverted \rf-relation and \co-relation: $\texttt{fr} = (\invrel{rf};\texttt{co})$. As an example memory model definition in \cat{} language, the the x86-TSO can be found in Appendix~\ref{apx:x86cat}. This memory model asserts acyclicity of communication relation, \texttt{po-loc} -relation, \texttt{mfence}-relation and some other derived relations~\cite{owens2009better}.

### 2018-05-15 3_encod.tex
Reduce space between letters in notations.

In general, a BMC problem aims to examine the non-reachability of the "undesirable" states of a finite-state system.
Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $Init(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $Trans(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $Bad(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps, is formulated as following:
%\begin{equation}
$\mathtt{SAT}( Init(\vec{x_0}) \land Trans(\vec{x}_0, \vec{x}_1) \land ... \land Trans(\vec{x}_{k-1}, \vec{x}_k) \land Bad(\vec{x}_k) )$.
%\end{equation}

@@

While working on the \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
For that, we compile the abstract syntax tree~(AST) of the parsed C-code to the plain event-flow graph.
We show %TODO! to show, actually!!!
that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner. This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions. 
For instance, \porthos{} used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} = (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\mathtt{if}\;x\;\mathtt{then}\;y\;\mathtt{else}\;z$').
In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula. The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:port:enc:df}).

@@

\begin{align}
\phi_{CF_{seq}}   \defeq \ & \ \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
\phi_{CF_{br}}    \defeq \ & \ [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \fx(e_1)] \land \nonumber \\
                      & \ [\fx(e_2) \rightarrow \fv(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
                      & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
\phi_{CF_{mer}} \defeq \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
\end{align}

@@

Following~\cite{Porthos17b}, the indexed references of variables are computed in accordance with the following rules:
%\begin{enumerate}[noitemsep,topsep=0pt]
(1)~any access to a shared variable (both read and write) increments its SSA-index;
(2)~only writes to a local variable increment its SSA-index (reads preserve indices);
(3)~no access to a constant variable or computed (evaluated) expression changes their %SSA-index.
%\end{enumerate}
These rules determine the following encoding of load, store and computation events within single thread:
%TODO: maybe add the overall formula \phi_{DF} ?
%
\begin{align}
    \phi_{DF_{e = \mathtt{load}(r \leftarrow l)}} \defeq \ & \fx(e) \rightarrow (r_{i+1} = l_{i+1}) \\
    \phi_{DF_{e = \mathtt{store}(l \leftarrow r)}} \defeq \ & \fx(e) \rightarrow (l_{i+1} = r_i) \\
    \phi_{DF_{e = \mathtt{eval}(\cdot)}} \defeq \ & \fx(e) \rightarrow \fv(e) % \\
\end{align}

@@

\begin{algorithm}
    \caption{Algorithm for computing the SSA-indices}\label{alg:compute-ssa}
    \algorithmicrequire{The event-flow graph $G = \langle N, E \rangle$ where $V$ is the set of nodes (events), $E$ is the set of control-flow transitions, $e_0$ is the entry node}
    \algorithmicensure{The SSA-map of the form "$\{$ event : $\{$ variable : index $\}\}$"}
    \begin{algorithmic}[1]
        \Function{Compute-SSA-Map}{G}
            \State {$S \gets$ empty map; $S[e_0] \gets$ empty map}
            \ForAll {event $e_i \in G.N$ in topological order}
                \ForAll {predecessor $e_j \in \mathtt{pred}(e_i)$}
                    \State {$S[e_i] \gets \mathtt{copy}(S[e_j])$}
                    \ForAll {variable $v_k \in$ set of variables accessed by $e_i$ }
                        \State {$S[e_i][v_k] \gets \mathtt{max}(S[e_i][v_k], \ S[e_j][v_k])$}
                        \If {need to update the index of $v_k$} \Comment {cases (1)-(2)}
                            \State {$S[e_i][v_k] \gets S[e_i][v_k] + 1$}
                        \EndIf
                    \EndFor
                \EndFor
            \EndFor
        \EndFunction
    \end{algorithmic}
    \label{algorithm:ssa-map}
\end{algorithm}

@@

As it has been described before, the \rf-relation links data-flow between events %TODO: another word?
of data-flow stored in equivalence assertions over the SSA-variables. 
%The encoding of the \rf-relation, which links the SSA-variables to the original variables, 
The encoding of this linkage left untouched as it is implemented in \porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
%
\begin{align}
\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \mathtt{rf}(e_1, e_2) \rightarrow (l_i = l_j) 
\end{align}

where the variable of location $l$ is mapped to the SSA-variable $l_i$ for event $e_1$, and to the SSA-variable $l_j$ for event $e_2$; and the predicate $\mathtt{rf}(e_1, e_2)$ is encoded as a boolean variable, which itself equals $true$ if $e_2$ reads the shared variable that was written in $e_1$.

@@

In an SMT-formula, the relation $x \relation{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
The derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
\begin{itemize}[noitemsep,topsep=0pt]
\item $r_1 \cup r_2(e_1, e_2) \defeq r_1(e_1, e_2) \lor r_2(e_1, e_2)$;
\item $r_1 \cap r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land r_2(e_1, e_2)$;
\item $r_1 \relminus r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land \neg r_2(e_1, e_2)$;
\item $r^{-1}(e_1, e_2) \defeq r(e_2, e_1)$;
\item $r^{*}(e_1, e_2) \defeq r^{+}(e_1, e_2) \lor (e_1 = e_2)$;
\item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb{E}}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$; \ and
\item $r^{+}(e_1, e_2) \defeq \mathtt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), \ where \\
\hspace*{1em} \mathtt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), \ and \\ 
\hspace*{1em} \mathtt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\mathtt{tc}_{i}(e_1,e_2);\mathtt{tc}_{i}(e_1,e_2))$.
\end{itemize}

Note that \cat{} language allows mutually-recursive definitions of relations (for example, `$\mathtt{r_1} = \mathtt{r_2} \cup (r_1;r_1)$').
The basic idea of using the Kleene fixpoint iteration for encoding such relations was also proposed in~\cite{Porthos17a}: for any pair of events $e_1,e_2 \in \mathbb{E}$ and relation $r \subseteq \mathbb{E} \times \mathbb{E}$, we encode a new integer variable $\Phi_{e_1,e_2}^{r}$ that represents the round of Kleene iteration on which the variable $r(e_1, e_2)$ has been set.
 %is used for guessing the execution

>>>>

In general, a BMC problem aims to examine the non-reachability of the "undesirable" states of a finite-state system.
Let $\vec{x} = (x_1, x_2, ..., x_n)$ be a vector of $n$ variables that uniquely distinguishes states of the system; let $\textit{Init}(\vec{x})$ be an \textit{initial-state predicate} that defines the set of initial states of the system; let $\textit{Trans}(\vec{x}, \pvec{x}')$ be a \textit{transition predicate} that signifies whether there the transition from state $\vec{x}$ to state $\pvec{x}'$ is valid; let $\textit{Bad}(\vec{x})$ be a \textit{bad-state predicate} that defines the set of undesirable states.
Then, the BMC problem, stated as the reachability of the undesirable state withing $k$ steps, is formulated as following:
%\begin{equation}
$\texttt{SAT}( \textit{Init}(\vec{x_0}) \land \textit{Trans}(\vec{x}_0, \vec{x}_1) \land ... \land \textit{Trans}(\vec{x}_{k-1}, \vec{x}_k) \land \textit{Bad}(\vec{x}_k) )$.
%\end{equation}

@@

While working on the \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
For that, we compile the \textit{abstract syntax tree~(AST)} of the parsed C-code to the plain event-flow graph.
We show %TODO! to show, actually!!!
that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner. This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions. 
For instance, \porthos{} used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
and control-flow of the branching instruction $i_1 = (c \; ? \; i_2 \; : \; i_3)$ was encoded as
$\phi_{CF}( \; c \; ? \; i_2 \; : \; i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula. The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:port:enc:df}).

@@

\begin{align}
\phi_{CF_{seq}}   \defeq \ & \ \fx(e_2) \Rightarrow \fx(e_1) \label{enc:cf:seq} \\
\phi_{CF_{br}}    \defeq \ & \ [\fx(e_2) \Rightarrow \fx(e_1)] \land [\fx(e_3) \Rightarrow \fx(e_1)] \land \nonumber \\
                      & \ [\fx(e_2) \Rightarrow \fv(e_1)] \land [\fx(e_3) \Rightarrow \lnot\fv(e_1)] \land \nonumber \\
                      & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
\phi_{CF_{mer}} \defeq \ & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \texttt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
\end{align}

@@

Following~\cite{Porthos17b}, the indexed references of variables are computed in accordance with the following rules:
%\begin{enumerate}[noitemsep,topsep=0pt]
(1)~any access to a shared variable (both read and write) increments its SSA-index;
(2)~only writes to a local variable increment its SSA-index (reads preserve indices);
(3)~no access to a constant variable or computed (evaluated) expression changes their %SSA-index.
%\end{enumerate}
These rules determine the following encoding of load, store and computation events within single thread:
%TODO: maybe add the overall formula \phi_{DF} ?
%
\begin{align}
    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}} \defeq \ & \fx(e) \Rightarrow (r_{i+1} = l_{i+1}) \\
    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & \fx(e) \Rightarrow (l_{i+1} = r_i) \\
    \phi_{DF_{e = \texttt{eval}(\cdot)}} \defeq \ & \fx(e) \Rightarrow \fv(e) % \\
\end{align}

@@

\begin{algorithm}
    \caption{Algorithm for computing the SSA-indices}\label{alg:compute-ssa}
    \algorithmicrequire{The event-flow graph $G = \langle N, E \rangle$ where $V$ is the set of nodes (events), $E$ is the set of control-flow transitions, $e_0$ is the entry node}
    \algorithmicensure{The SSA-map of the form "$\{$ event : $\{$ variable : index $\}\}$"}
    \begin{algorithmic}[1]
        \Function{Compute-SSA-Map}{G}
            \State {$S \gets$ empty map; $S[e_0] \gets$ empty map}
            \ForAll {event $e_i \in G.N$ in topological order}
                \ForAll {predecessor $e_j \in \texttt{pred}(e_i)$}
                    \State {$S[e_i] \gets \texttt{copy}(S[e_j])$}
                    \ForAll {variable $v_k \in$ set of variables accessed by $e_i$ }
                        \State {$S[e_i][v_k] \gets \texttt{max}(S[e_i][v_k], \ S[e_j][v_k])$}
                        \If {need to update the index of $v_k$} \Comment {cases (1)-(2)}
                            \State {$S[e_i][v_k] \gets S[e_i][v_k] + 1$}
                        \EndIf
                    \EndFor
                \EndFor
            \EndFor
        \EndFunction
    \end{algorithmic}
    \label{algorithm:ssa-map}
\end{algorithm}

@@

As it has been described before, the \rf-relation links data-flow between events %TODO: another word?
of data-flow stored in equivalence assertions over the SSA-variables. 
%The encoding of the \rf-relation, which links the SSA-variables to the original variables, 
The encoding of this linkage left untouched as it is implemented in \porthos: for each pair of events $e_1$ and $e_2$ linked by the \rf-relation, we add the following constraint:
%
\begin{align}
\phi_{DF_{mem}}(e_1, e_2) \defeq \ & \texttt{rf}(e_1, e_2) \rightarrow (l_i = l_j)
\end{align}

where the variable of location $l$ is mapped to the SSA-variable $l_i$ for event $e_1$, and to the SSA-variable $l_j$ for event $e_2$; and the predicate $\texttt{rf}(e_1, e_2)$ is encoded as a boolean variable, which itself equals $true$ if $e_2$ reads the shared variable that was written in $e_1$.

@@

In an SMT-formula, the relation $x \relarrow{r} y$ is represented by a boolean variable $r(x, y)$ that indicates whether the relation holds.
The derived relations are encoded by fresh boolean variables according to the following rules~\cite{Porthos17b}:
\begin{itemize}[noitemsep,topsep=0pt]
\item $r_1 \cup r_2(e_1, e_2) \defeq r_1(e_1, e_2) \lor r_2(e_1, e_2)$;
\item $r_1 \cap r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land r_2(e_1, e_2)$;
\item $r_1 \relminus r_2(e_1, e_2) \defeq r_1(e_1, e_2) \land \neg r_2(e_1, e_2)$;
\item $r^{-1}(e_1, e_2) \defeq r(e_2, e_1)$;
\item $r^{*}(e_1, e_2) \defeq r^{+}(e_1, e_2) \lor (e_1 = e_2)$;
\item $r_1;r_2(e_1,e_2) \defeq \bigvee\limits_{e_k \in \mathbb{E}}^{} r_1(e_1,e_k) \land r_2(e_k,e_2)$; \ and
\item $r^{+}(e_1, e_2) \defeq \texttt{tc}_{\left\lceil log|\mathbb{E}| \right\rceil} (e_1, e_2), \ where \vspace{0.25em} \\
\texttt{tc}_{0}(e_1,e_2) \defeq r(e_1,e_2), \ and \\
\texttt{tc}_{i+1}(e_1,e_2) \defeq r(e_1,e_2) \lor (\texttt{tc}_{i}(e_1,e_2);\texttt{tc}_{i}(e_1,e_2))$.
\end{itemize}
\vspace{1em}

Note that \cat{} language allows mutually-recursive definitions of relations (for example, `$\texttt{r}_\texttt{1} = \texttt{r}_\texttt{2} \cup (\texttt{r}_\texttt{1}; \texttt{r}_\texttt{1})$').
The basic idea of using the Kleene fixpoint iteration for encoding such relations was also proposed in~\cite{Porthos17a}: for any pair of events $e_1,e_2 \in \mathbb{E}$ and relation $r \subseteq \mathbb{E} \times \mathbb{E}$, we encode a new integer variable $\Phi_{e_1,e_2}^{r}$ that represents the round of Kleene iteration on which the variable $r(e_1, e_2)$ has been set.
 %is used for guessing the execution
